### 2015-10-13 manuscript/conclusions.tex
Restructured paper directory for conference submission.

There are still many other factors to consider in forming a better models. From this research, we found that the accurate prediction of drunkenness in a real application looks possible in theory. There are still many obstacles surrounding the collection of a larger and better data set. Ideally, we would want a data set from a thousand volunteers in candid situations over several days. The two biggest problems are, how do we label the data with the alcohol levels, and how can we model this enormous amount of data in reasonable time? Also, will we need to first determine the activity of the user, or will the other sensor provide sufficient information? And if the former, how can we consistently tag these activities in the data? The answers to this may be simple, or infeasible. In any case, it would be worth it to try and find out.

>>>>

There are still many other factors to consider in forming a better models. From this research, we found that the accurate prediction of drunkenness in a real application looks possible in theory. There are still many obstacles surrounding the collection of a larger and better data set. Ideally, we would want a data set from a thousand volunteers in candid situations over several days. The two biggest problems are, how do we label the data with the alcohol levels in this scenario, and how can we model this enormous amount of data in reasonable time? Also, will we need to first determine the activity of the user, or will the other sensors provide sufficient information? And if the former, how can we consistently tag these activities in the data? The answers to this may be simple, or infeasible. In any case, it would be worth it to try and find out.

### 2015-10-13 manuscript/intro.tex
Restructured paper directory for conference submission.

\section{Introduction}
%%Describe the problem, state your contributions and the organization of the paper.T

Internet of Things (IoT) is a domain that represents the next most exciting technological revolution since the Internet. IoT will bring endless opportunities and impact every corner of our planet. In the healthcare domain, IoT promises to bring personalized health tracking and monitoring ever closer to the consumers. This phenomena is evidenced in a recent Wall Street Journal (June, 29, 2015) article entitled "Staying Connected is Crucial to Staying Healthy". Modern smartphones and related devices now contain more sensors than ever before. Data from sensors can be collected more easily and more accurately. In 2014, it is estimated that 46 million people are using IoT-based health and fitness applications. Currently, the predominant IoT-based health applications are in sports and fitness. However, disease management or preventive care health applications are becoming more prevalent. For example, the real-time preventive care applications such as those for detecting fall in elderly patients is one of the active research areas due to the aging population \cite{Tacconi:2011}. This paper describes a new category of real-time preventive health application that aims to reduce risky behaviors due to intoxication by predicting blood alcohol content (BAC) non-invasively and in real-time.

The Centers for Disease Control and Prevention (CDC) reported \cite{Schoenborn:2013} that between the years 2008 and 2010 roughly two-thirds of adults were current drinkers, but adults between the ages 18 and 24 were most commonly associated with consumption of more than 5 drinks in one day. There is also a known causal relationship between alcohol consumption and risky behaviors \cite{Assaad:2006}. One of the worst risky behaviors from drinking is drunk driving which is a persistent problem worldwide. Innocent people are harmed/killed in accidents where the driver was under the influence. It is reported by the Bureau of Transportation Statistics \cite{Chambers:2012} that in 2010 47.2\% of pedestrian fatalities and 39.9\% of vehicle occupant fatalities are caused by drunk driving.

Being able to calculate BAC is important for the portion of the population that consumes alcohol. Drinkers often find themselves in situations in which they are drinking but also need to plan on driving home. The issue comes when the drinker is not sure if they are sober enough, as in their BAC is below the legal limit, to drive home safely. In many cases, drinkers misjudge their intoxication, believing they will be able to safely get home, when actually their BAC is higher than the legal limit. While there are some ways available for people to measure their BAC, these methods are often invasive, inaccurate, or impractical.

The most commonly known method of measuring BAC is the use of a breathalyzer. These devices are portable, easily accessible, and accurate. However, breathalyzers have a stigma of DUI arrest, fines, and license suspension. Breathalyzers are also non-discreet and invasive because the device requires the user to blow into the device in order to measure BAC. Due to these reasons, people often do not want to use such a device. A noninvasive, discreet, and real-time method of measuring BAC is desired.

In this paper, we propose an android-based IoT middleware architecture and application for collecting sensor data on smartwatches and smartphones to predict BAC in real-time. A study conducted in Washington University \cite{Ward:2015} shows that electronic intervention programs are more successful at reducing college student drinking than general alcohol awareness program. By wearing a smartwatch, a user can be alerted in real time when the BAC is above certain level and programmed with different options of intervention when such an event occurs. For example, the application can send a text message to a trusted family member or friend or can lock the ignition of the car engine of the drinker. Currently, there are no known applications that leverages data from common wearable devices to predict BAC non-invasively and in real time.

The main contributions of the paper are:
\begin{itemize}
	\item A middleware platform that can collect data from any type of device (virtual or physical) that can communicate with android phones. 
	\item A BAC prediction model based on live smartwatch sensor data that enables users to get BAC measurement in real time.
	\item Geographical visualization of BAC predictions.
	\item Collection of a labeled BAC data set for training the prediction model.
\end{itemize}

%****Summary of the results and conclusion****

>>>>

\section{Introduction}
Internet of Things (IoT) is a domain that represents the next most exciting technological revolution since the Internet. IoT will bring endless opportunities and impact every corner of our planet. In the healthcare domain, IoT promises to bring personalized health tracking and monitoring ever closer to the consumers. This phenomena is discussed in a recent Wall Street Journal article, "Why Connected Medicine Is Becoming Vital to Health Care" \cite{Landro:2015}. Modern smartphones and smartwatches now contain a more diverse collection of sensors than ever before, and people are warming up to them. In January 2014, approximately 46 million US smartphone owners were reported to have used health and fitness applications \cite{Nielsen:2014}. Currently, sports and fitness are the predominant foci of IoT-based health applications. However, applications in disease management and health care are becoming increasingly prevalent. For example, detecting falling of elderly patients \cite{Tacconi:2011}.

Drunk driving is a dangerous, worldwide problem. This problem is not only a hazard to the drunk drivers, but also to pedestrians and other drivers. It is reported by the Bureau of Transportation Statistics that in 2010, 47.2\% of pedestrian fatalities and 39.9\% of vehicle occupant fatalities were caused by drunk driving \cite{Chambers:2012}. The Centers for Disease Control and Prevention (CDC) reported that between the years 2008 and 2010, roughly two-thirds of adults were drinkers, with adults between the ages 18 and 24 having the greatest association with heavy drinking \cite{Schoenborn:2013}.

At dangerous levels of intoxication, it can be difficult to judge ones own drunkenness. Instead it would be better to get a definitive measurement of the BAC, or simply a binary response: "drunk" or "not drunk." Compact breathalyzers are probably the best option at the moment, but these are not discreet and require deliberate action by the user. The other option is to use a smartphone application to manually calculate BAC, but these demand a greater deal of involvement from the user. To be practical, it would be useful to have some sort of non-invasive and accurate monitoring system that will warn its user if they become too intoxicated. It has been shown that electronic intervention programs are more successful at reducing college student drinking than a general alcohol awareness program \cite{Ward:2015}. This system can also be used to warn friends and family, or prevent the operation of the user's car.

In this paper, we investigate the prediction of intoxication level from smartwatch sensor data via machine learning. We also briefly discuss a general Android-based gateway system which can collect data from any type of physical or virtual sensor accessible by the host smartphone.

### 2015-10-13 manuscript/system.tex
Restructured paper directory for conference submission.

\section{System Architecture}
A custom Android application (Figure \ref{app}) was designed to be used during the study. This application collected and locally stored sensor data from a Microsoft Band smartwatch in a CSV file. The following data was collected from the Microsoft Band:
\begin{itemize}
	\item Accelerometer
	\item Gyroscope
	\item Pace
	\item Speed
	\item Distance traveled
	\item Motion type (idle, walking, etc.)
	\item Heart rate
	\item Pedometer
	\item Skin temperature
	\item Ultraviolet level
	\item Contact (if smartwatch is on)
	\item Calories burned
\end{itemize}
Additional data continuously collected and stored in the CSV by the application included:
\begin{itemize}
	\item GPS coordinates from the smartphone 
	\item Button that was pressed whenever the user took a drink
	\item Input field used to periodically record the BAC measured by a separate breathalyzer.
\end{itemize}

\begin{figure}[H]
	\centering
	\includegraphics[scale=.20]{smartwatch.png}
	\caption{Capture of the Android application}
	\label{app}
\end{figure}

User information was also recorded and stored through the application before data collection began. Each user was given an unique user id (UUID) and the user profile included age, sex, body mass index, and blood type which was also stored locally as a JSON file.

After the data had been collected, the stored csv file and json file were uploaded to the web server and data inserted into mySQl database by a php script. Data stored on the MySQL database was then displayed on a web client by a geographical map.

\begin{figure}[H]
	\centering
	\includegraphics[scale=.35]{webpage.png}
	\caption{Current data visualization}
	\label{webpage}
\end{figure}

As of now, the web client displays the last sample for each user at the recorded location (Figure \ref{webpage}). Ideally in future systems, many samples will be averaged across regions (Figure \ref{concept}).

\begin{figure}[H]
	\centering
	\includegraphics[scale=.23]{webconcept.png}
	\caption{Ideal data visualization}
	\label{concept}
\end{figure}

The collected data was also used to train a machine learning model which would be implemented in a future Android application to predict user BAC or drunkenness.

\begin{figure}[H]
	\centering
	\includegraphics[scale=.30]{applicationsystem.png}
	\caption{Application system}
	\label{appsystem}
\end{figure}

Representation of this system can be found in Figure \ref{overallsystem}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=.14]{overall_system.png}
	\caption{Overall system}
	\label{overallsystem}
\end{figure}

>>>>

\section{System Architecture}

A huge end-goal of ours is to have a public system where participating users' phones transmit sensor information to a central database through a REST-based web service. The service stores this information in the database and another data processing system. Other users who are not contributing information can connect to our system to access other related services, like a choropleth map of the sensor data; BAC prediction data in our case, but the system would not be limited to it. The particular protocols and setup used should be designed to protect the privacy of the sensor data contributors.

\begin{figure}
	\includegraphics[width=1.0\textwidth]{figs/system}
	\caption{Example general system for large-scale sensor data collection}
	\label{fig:system}
\end{figure}

\subsection{Smartphone Sensor Gateway}

Each of the smartphones in this system are behaving as a type of sensor gateway. In order to collect our data, we built such a system on the Android platform. Though we used primarily a Microsoft Band, our system allows the easy addition of any sensor that can be connected to from the Android smartphone. In Figure \ref{fig:gateway}, we show an overview of the classes used in this system.

\begin{figure}
	\includegraphics[width=1.0\textwidth]{figs/gateway}
	\caption{Gateway implementation for our data collection application}
	\label{fig:gateway}
\end{figure}
